---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
devtools::load_all()
```

# tibbletime 

[![Travis-CI Build Status](https://travis-ci.org/business-science/tibbletime.svg?branch=master)](https://travis-ci.org/business-science/tibbletime.svg?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tibbletime)](https://cran.r-project.org/package=tibbletime)

## A time aware tibble

Built on top of the `tidyverse`, `tibbletime` is an extension
that allows for the creation of _time aware tibbles_ through the setting of 
a time index. 

Some immediate advantages of this include: 

1) The ability to perform compact time based subsetting on tibbles.

2) Quickly summarising and aggregating results by time period
(yearly, monthly, etc).

3) Changing the periodicity of a time based tibble. This means changing 
from a daily dataset to a monthly or yearly dataset.

4) Calling functions similar in spirit to the `map` family from `purrr` 
on time based tibbles.

5) All functions were designed to support the pipe and to work with packages
like `dplyr` and `tidyr`.

## Installation

Development Version:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("business-science/tibbletime")
```

CRAN Version: 

```{r, eval = FALSE}
install.packages("tibbletime")
```

## Getting started

The first thing to do is to turn your `tibble` into a `tbl_time` object.
Notice the specification of the `index` as the `date` column of `FB`.

```{r, message=FALSE, warning=FALSE}
library(tibbletime)

# Facebook stock prices. Comes with the package
data(FB)

# Convert FB to tbl_time
FB <- FB %>% as_tbl_time(index = date)

FB
```

There are a number of functions that were designed specifically for `tbl_time`
objects. Some of them are:

1) `time_filter` - Succinctly filter a tbl_time object by date.

2) `time_summarise` - Similar to dplyr::summarise but with the added benefit 
of being able to summarise by a time period such as "yearly" or "monthly".

3) `tmap` - The family of tmap functions transform a tbl_time input by applying a 
function to each column at a specified time interval.

4) `as_period` - Convert a tbl_time object from daily to monthly, 
from minute data to hourly, and more. This allows the user to easily 
aggregate data to a less granular level.

5) `time_collapse` - When time_collapse is used, the index of a tbl_time object 
is altered so that all dates that fall in a period share a common date.

To look at just a few:

```{r}
# Filter for dates from March 2013 to December 2015
FB %>% 
  time_filter(2013-03 ~ 2015)

# Change from daily to monthly periodicity
FB %>% 
  as_period("monthly")

# Get the average mean and standard deviation for each year
FB %>%
  time_summarise(period = "yearly",
        adj_mean = mean(adjusted),
        adj_sd   = sd(adjusted))
```

## Grouping

Groups created through `dplyr::group_by()` are supported throughout the package.

```{r}
# Facebook, Amazon, Netflix and Google stocks
data(FANG)

# Summarise by period and by group
FANG %>% 
  as_tbl_time(date) %>%
  group_by(symbol) %>%
  time_summarise(period = "yearly",
        adj_min   = min(adjusted),
        adj_max   = max(adjusted),
        adj_range = adj_max - adj_min)
```


## Warning

This package is still going through active development and is subject to change. Use at your own risk. Reproducible bug reports are welcome.

Vignettes and more functionality are coming soon. Stay tuned.
