---
title: "Time-based filtering"
author: "Davis Vaughan"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Time-based filtering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tibbletime)
```

## `time_filter()`

`time_filter()` attempts to make filtering data frames by date much easier 
than `dplyr::filter()`. It includes a flexible shorthand notation that allows
you to specify entire date ranges with very little typing. The general form 
of the `time_formula` that you will use to filter rows is `from ~ to`, where 
the left hand side (LHS) is the start date, and the right hand side (RHS) is
the end date. Both endpoints are included. Each side of the `time_formula` can
be maximally specified as `YYYY-MM-DD + HH:MM:SS`.

### Datasets required

```{r, message=FALSE, warning=FALSE}
library(tibbletime)

# Facebook stock prices.
data(FB)

# Convert FB to tbl_time
FB <- as_tbl_time(FB, index = date)

# FANG stock prices
data(FANG)

# Convert FANG to tbl_time and group
FANG <- as_tbl_time(FANG, index = date) %>%
  group_by(symbol)

```

### Year filtering example

In `dplyr`, if you wanted to get the dates for `2013` in the `FB` dataset, you 
might do something like this:

```{r}
filter(FB, date >= as.Date("2013-01-01"), date <= as.Date("2013-12-31"))
```

That's a lot of typing for one filter step. With `tibbletime`, because
the `index` was specified at creation, we can do this:

```{r}
time_filter(FB, time_formula = 2013-01-01 ~ 2013-12-31)
```

At first glance, this might not look like less code, but this is before any
shorthand is applied. Note how the filtering condition is specified as a 
`formula` separated by a `~`. 

Using `time_filter` shorthand, this can be written:

```{r}
time_filter(FB, 2013 ~ 2013)
```

Or even more succinctly as:

```{r}
time_filter(FB, ~2013)
```

The shorthand notation works as follows. In the first example, `2013 ~ 2013` is
expanded to `2013-01-01 + 00:00:00 ~ 2013-12-31 + 23:59:59`. It works by 
identifying the periodicity of the provided input (yearly), and expanding it
the the beginning and end of that period. The one sided formula `~2013` works 
similarly, and it useful when you want to select every date inside a period. 

### Month filtering example

As another example of this shorthand, if you wanted to select every date in
March, 2015:

```{r}
time_filter(FB, ~2015-03)

# In dplyr it looks like this
# (and you have to think, does March have 30 or 31 days?)
# filter(FB, date >= as.Date("2015-03-01"), date <= as.Date("2015-03-31"))
```

### Grouped example

Working with grouped `tbl_time` objects is just as you might expect.

```{r}
FANG %>%
  time_filter(2013-01-01 ~ 2013-01-04)
```

### Finer periods

Filtering can also be done by hour / minute / second. Note that the form of this
is slightly different than the standard, `YYYY-MM-DD + HH:MM:SS`. The big 
difference here is that a `+` is required to divide the date from the time.

```{r}
# Dummy example. Every second in a day
example <- create_series(~2013-01-01, period = 1~s)

# The first 2 minutes of the day
example %>%
  time_filter(2013-01-01 ~ 2013-01-01 + 00:02)

# 3 specific hours of the day
# Equivalent to:
# 2013-01-01 + 03:00:00 ~ 2013-01-01 + 06:59:59
example %>%
  time_filter(2013-01-01 + 3 ~ 2013-01-01 + 6)
```

### `[` syntax

For interactive use, to get an even quicker look at a dataset you can use 
the traditional extraction operator `[` with the formula syntax.

```{r}
FB[~2013]
```

```{r}
FB[2013~2014-02, c(1,2,3)]
```

