---
title: "Time-based filtering"
author: "Davis Vaughan"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Time-based filtering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introducing filter_time()

`filter_time()` attempts to make filtering data frames by date much easier 
than `dplyr::filter()`. It includes a flexible shorthand notation that allows
you to specify entire date ranges with very little typing. The general form 
of the `time_formula` that you will use to filter rows is `from ~ to`, where 
the left hand side (LHS) is the character start date, and the right hand side (RHS) is
the character end date. Both endpoints are included in the result.
Each side of the `time_formula` can be maximally specified as 
the character `'YYYY-MM-DD HH:MM:SS'`.

## Datasets required

```{r, message=FALSE, warning=FALSE}
library(tibbletime)
library(dplyr)

# Facebook stock prices.
data(FB)

# Convert FB to tbl_time
FB <- as_tbl_time(FB, index = date)

# FANG stock prices
data(FANG)

# Convert FANG to tbl_time and group
FANG <- as_tbl_time(FANG, index = date) %>%
  group_by(symbol)

```

## Year filtering example

In `dplyr`, if you wanted to get the dates for `2013` in the `FB` dataset, you 
might do something like this:

```{r}
filter(FB, date >= as.Date("2013-01-01"), date <= as.Date("2013-12-31"))
```

That's a lot of typing for one filter step. With `tibbletime`, because
the `index` was specified at creation, we can do this:

```{r}
filter_time(FB, time_formula = '2013-01-01' ~ '2013-12-31')
```

At first glance, this might not look like less code, but this is before any
shorthand is applied. Note how the filtering condition is specified as a 
`formula` separated by a `~`. 

Using `filter_time` shorthand, this can be written:

```{r}
filter_time(FB, '2013' ~ '2013')
```

Or even more succinctly as:

```{r}
filter_time(FB, ~'2013')
```

The shorthand notation works as follows. In the first example, `'2013' ~ '2013'` is
expanded to `'2013-01-01 + 00:00:00' ~ '2013-12-31 + 23:59:59'`. It works by 
identifying the periodicity of the provided input (yearly), and expanding it
to the beginning and end of that period. The one sided formula `~'2013'` works 
similarly, and is useful when you want to select every date inside a period. 

## Month filtering example

As another example of this shorthand, if you wanted to select every date in
March, 2015:

```{r}
filter_time(FB, ~'2015-03')

# In dplyr it looks like this
# (and you have to think, does March have 30 or 31 days?)
# filter(FB, date >= as.Date("2015-03-01"), date <= as.Date("2015-03-31"))
```

## Keywords

Two keywords are available to assist with filtering:

* `'start'` - The start of the series
* `'end'` - The end of the series

This filters from the start of the series to the end of 2015.

```{r}
filter_time(FB, 'start' ~ '2015')
```


## Grouped example

Working with grouped `tbl_time` objects is just as you might expect.

```{r}
FANG %>%
  filter_time('2013-01-01' ~ '2013-01-04')
```

## Finer periods

Filtering can also be done by hour / minute / second. Note that the form of this
is slightly different than the standard, `'YYYY-MM-DD HH:MM:SS'`.

```{r}
# Dummy example. Every second in a day
example <- create_series(~'2013-01-01', period = 's')

# The first 2 minutes of the day
example %>%
  filter_time('2013-01-01' ~ '2013-01-01 00:02')

# 3 specific hours of the day
# Equivalent to:
# '2013-01-01 + 03:00:00' ~ '2013-01-01 + 06:59:59'
example %>%
  filter_time('2013-01-01 3' ~ '2013-01-01 6')
```

## `[` syntax

For interactive use, to get an even quicker look at a dataset you can use 
the traditional extraction operator `[` with the formula syntax.

```{r}
FB[~'2013']
```

```{r}
FB['2013'~'2014-02', c(1,2,3)]
```

## Using variables in the filter

Each side of the time formula is unquoted and evaluated in the environment 
that is was created using `rlang`. This means that you can use variables inside
the call the `filter_time()`.

```{r}
date_var <- as.Date("2014-01-01")
filter_time(FB, 'start' ~ date_var)

date_char <- "2014-02"
filter_time(FB, ~ date_char)
```

