---
title: "Introduction to tibbletime"
author: "Davis Vaughan"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction to tibbletime}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tibbletime)
```

## A time aware tibble

Built on top of the `tidyverse`, `tibbletime` is an extension
that allows for the creation of _time aware tibbles_ through the setting of 
a time index. 

Some immediate advantages of this include: 

1) The ability to perform compact time-based subsetting on tibbles.

2) Quickly summarising and aggregating results by time period
(yearly, monthly, every 2 weeks, etc).

3) Changing the periodicity of a time-based tibble. This means changing 
from a daily dataset to a monthly or yearly dataset.

4) Calling functions similar in spirit to the `map()` family from `purrr` 
on time-based tibbles.

5) All functions were designed to support the pipe and to work with packages
like `dplyr` and `tidyr`. Each function has also been designed to work with
`dplyr::group_by()` allowing for powerful data manipulation.

6) Modifying functions for rolling analysis.

7) Quickly creating `tbl_time` time series objects.

## Getting started

The first thing to do is to turn your `tibble` into a `tbl_time` object.
Notice the specification of the `index` as the `date` column of `FB`.

```{r, message=FALSE, warning=FALSE}
library(tibbletime)

# Facebook stock prices. Comes with the package
data(FB)

# Convert FB to tbl_time
FB <- as_tbl_time(FB, index = date)

FB
```

There are a number of functions that were designed specifically for `tbl_time`
objects. Some of them are:

1) `time_filter` - Succinctly filter a tbl_time object by date.

2) `time_summarise` - Similar to dplyr::summarise but with the added benefit 
of being able to summarise by a time period such as "yearly" or "monthly".

3) `tmap` - The family of tmap functions transform a tbl_time input by applying a 
function to each column at a specified time interval.

4) `as_period` - Convert a tbl_time object from daily to monthly, 
from minute data to hourly, and more. This allows the user to easily 
aggregate data to a less granular level.

5) `time_collapse` - When time_collapse is used, the index of a tbl_time object 
is altered so that all dates that fall in a period share a common date.

6) `rollify()` - Modify a function so that it calculates a value 
(or a set of values) at specific time intervals. This can be used for rolling
averages and other rolling calculations inside the `tidyverse` framework.

7) `create_series()` - Use shorthand notation to quickly 
initialize a `tbl_time` object containing a `date` column 
with a regularly spaced time series.

To look at just a few:

```{r}
# Filter for dates from March 2013 to December 2015
FB %>% 
  time_filter(2013-03 ~ 2015)

# Change from daily to monthly periodicity
FB %>% 
  as_period("monthly")

# Get the average mean and standard deviation for each year
FB %>%
  time_summarise(period = 1~y,
        adj_mean = mean(adjusted),
        adj_sd   = sd(adjusted))

# Perform a 5 period rolling average
mean_5 <- rollify(mean, window = 5)
FB %>%
  mutate(roll_mean = mean_5(adjusted))

# Create a time series
# Every other day in 2013
create_series(~2013, 2~d)
```

## Grouping

Groups created through `dplyr::group_by()` are supported throughout the package.

```{r}
# Facebook, Amazon, Netflix and Google stocks
data(FANG)

# Summarise by period and by group
FANG %>% 
  as_tbl_time(date) %>%
  group_by(symbol) %>%
  time_summarise(period = "yearly",
        adj_min   = min(adjusted),
        adj_max   = max(adjusted),
        adj_range = adj_max - adj_min)
```
